A quantum parallel Markov chain Monte Carlo

### 2022-03-03 main.tex
moving toward final exp
new results and more writing

\subsection{QPMCMC}

\begin{algorithm}[t!]
	\caption{Quantum parallel MCMC}\label{alg:qpMCMC}
	\KwData{Initial Markov chain state $\ttheta^{(0)}$; total length of Markov chain $S$; total number of proposals per iteration $P$; routine for evaluating target density $\pi(\cdot)$; routines for drawing random samples from the proposal distribution $Q(\ttheta^{(s)},\cdot)$ and the standard Gumbel distribution $Gumbel(0,1)$.}
	\KwResult{A Markov chain $\ttheta^{(1)}, \dots, \ttheta^{(S)}$.}
	\For{$s \in\{1,\dots,S\}$}{
		%$\sigma_\pi \gets 0$\;
		$\ttheta_0 \gets \ttheta^{(s-1)}$\;
		\For{$p' \in \{1,\dots,P\}$}{
			$\ttheta_{p'} \gets Q(\ttheta^{(s-1)},\cdot)$\;
		}
		\For{$p' \in \{0,\dots,P\}$}{
			$q_{p'} \gets  \prod_{p''\neq p'}q(\ttheta_{p'},\ttheta_{p''})$\; 
			$z_{p'} \gets Gumbel(0,1)$\;
		}
		$p \gets \argmin_{p'=0,\dots,P}f(p')= -\left(\log \big(\pi(\ttheta_{p'}) q_{p'}\big) + z_{p'}\right)$; \hspace{2em} \Comment{$\order{\sqrt{P}}$ Alg \ref{alg:min}}
		$\ttheta^{(s)} \gets \ttheta_p$\;
	}
	\Return{$\ttheta^{(1)}, \dots, \ttheta^{(S)}$}\ .

@@

Algorithm \ref{alg:qpMCMC} presents the details of QPMCMC.  The algorithm uses conventional (perhaps parallel) computing almost entirely, excluding the one line that calls the quantum minimization algorithm presented in Algorithm \ref{alg:min}. This sub-routine seeks to find the minimizer of the function
\begin{align*}
f(p')= -\left(\log \big(\pi(\ttheta_{p'}) q_{p'}\big) + z_{p'}\right) \quad \mbox{over} \quad p'=0,1,\dots, P,
\end{align*} 
which combines the numerators of Equation \eqref{eq:probs}, the Gumbel-max trick and a trivial negation.  As discussed in Section \ref{sec:qmin}, the sub-routine requires only $\mathcal{O}(\sqrt{P})$ oracle evaluations and, therefore, only $\mathcal{O}(\sqrt{P})$ evaluations of the target density $\pi(\cdot)$ using a quantum circuit.  In theory, a quantum computer can perform the same computations as a classical computer, but the efficiency of the target evaluations would depend on a number of factors related to the structure of the density itself, the availability of fast quantum random access memory \citep{giovannetti2008quantum} and the ability of large numbers of quantum gates to act in concert with negligible noise \citep{kielpinski2002architecture,erhard2019characterizing}.

@@

>>>>

\subsection{Symmetric proposals}

\subsection{QPMCMC}
\newcommand{\SSigma}{\boldsymbol{\Sigma}}

\begin{algorithm}[t!]
	\caption{Quantum parallel MCMC}\label{alg:qpMCMC}
	\KwData{Initial Markov chain state $\ttheta^{(0)}$; total length of Markov chain $S$; total number of proposals per iteration $P$; routine for evaluating target density $\pi(\cdot)$; routines for drawing random samples from a $D$-dimensional multivariate Gaussian $Normal_D(\boldsymbol{\mu},\SSigma)$ and the standard Gumbel distribution $Gumbel(0,1)$.}
	\KwResult{A Markov chain $\ttheta^{(1)}, \dots, \ttheta^{(S)}$.}
	\For{$s \in\{1,\dots,S\}$}{
		$\ttheta_0 \gets \ttheta^{(s-1)}$\;
		$\bar{\ttheta} \gets Normal_D(\ttheta_0,\SSigma)$; \hspace{15em} \Comment{ \hspace{0.3em}     Proposal \eqref{alg:min}}      
		\For{$p \in \{0,\dots,P\}$}{
			$\ttheta_p \gets Normal_D(\bar{\ttheta},\SSigma)$\;
			$z_{p} \gets Gumbel(0,1)$\;
		}
		$\hat{p} \gets \argmin_{p=0,\dots,P}\Big(f(p):= -\big(z_{p}+ \log\pi(\ttheta_{p}) \big)\Big)$; \hspace{3.5em} \Comment{$\order{\sqrt{P}}$ Alg \ref{alg:min}}
		$\ttheta^{(s)} \gets \ttheta_{\hat{p}}$\;
	}
	\Return{$\ttheta^{(1)}, \dots, \ttheta^{(S)}$}\ .

@@

Algorithm \ref{alg:qpMCMC} presents the details of QPMCMC.  The algorithm uses conventional (perhaps parallel) computing almost entirely, excluding the one line that calls the quantum minimization algorithm presented in Algorithm \ref{alg:min}.  To keep the conventional computing burden low, we recommend

This sub-routine seeks to find the minimizer of the function
\begin{align*}
f(p)= -\big(z_{p}+ \log\pi(\ttheta_{p}) \big) \quad \mbox{over} \quad p=0,1,\dots, P,
\end{align*} 
which combines the numerators of Equation \eqref{eq:probs}, the Gumbel-max trick and a trivial negation.  As discussed in Section \ref{sec:qmin}, the sub-routine requires only $\mathcal{O}(\sqrt{P})$ oracle evaluations and, therefore, only $\mathcal{O}(\sqrt{P})$ evaluations of the target density $\pi(\cdot)$ using a quantum circuit.  In theory, a quantum computer can perform the same computations as a classical computer, but the efficiency of the target evaluations would depend on a number of factors related to the structure of the density itself, the availability of fast quantum random access memory \citep{giovannetti2008quantum} and the ability of large numbers of quantum gates to act in concert with negligible noise \citep{kielpinski2002architecture,erhard2019characterizing}.

@@

%\begin{table}[ht]\centering
%	\begin{tabular}{llll}  \toprule 
%	Proposals	& Effective samples (ES) & Evaluations per ES & Speedup \\   \midrule
%	1 (MH)	 & 21.35  (11.74, 50.65) & 429.81  (180.15, 766.89) & 1.00  (1.00, 1.00) \\  
%		 50 & 736.25 (418.73, 1072.35) & 259.54  (177.72, 475.16) & 1.66  (1.01, 1.61) \\  
%		 100 & 1842.71 (1604.07, 2097.74) & 146.49  (128.76, 168.39) & 2.93 (1.40, 4.55) \\  
%		 150 & 2287.03  (1734.34, 2444.05) & 146.39 (137.36, 193.16) & 2.94 (1.31, 3.97) \\ 
%		  200 & 2662.89  (2547.71, 2844.14) & 145.92 (136.37, 152.74) & 2.95  (1.32, 5.02) \\ 
%		   250 & 2971.59  (2439.45, 3208.46) & 147.77  (136.53, 180.95) & 2.91  (1.32, 4.24) \\  
%		  \bottomrule\end{tabular}
%\end{table}
